---
title: Interactive visualizations in R
runtime: shiny
output: html_document
author:
- name: Sebastian Navdahl,
- name: Johannes Kolberg Krokeide,
- name: Benjamin Aanes,
- name: Marcus Lauritsen
  affiliation: Norwegian School of Economics
---

# 1. Introduction
- Provides the possibility to dig deeper into the data. 

# 2. Housing prices dashboard
Introduction.

```{r echo=TRUE, results='hide'}
library(shiny)
library(DT)
library(dplyr)
library(readr)

# Load data
house <- read.csv('houseprices.csv')

# Get lists of all neighbourhoods and which to exclude by default, for plotting
house$Neighborhood = as.factor(house$Neighborhood) # for levels to be pulled, it needs to be formatted as factor
all_neighbourhoods <- levels(pull(house, 'Neighborhood'))
exclude_neighbourhoods <- setdiff(levels(house$Neighborhood), c("Blmngtn", "StoneBr", "Mitchel", "OldTown", "NoRidge"))

# Define UI for application
fluidPage(
  # Application title
  titlePanel("Iowa Housing: Sale Price"),
  
  # Sidebar with various Exploratory Data Analysis choices
  sidebarLayout(
    sidebarPanel(
       # Select variable for x-axis
       selectInput(inputId = "x",
                   label = "X-axis:",
                   choices = c("GrLivArea", "TotalSF", "TotalBsmtSF", "HouseAge"),
                   selected = "GrLivArea"),
       
       # Select variable for colour
       selectInput(inputId = "z",
                   label = "Colour:",
                   choices = c("Neighborhood", "CentralAir", "MSSubClass", "SaleType", "SaleCondition", "Foundation"),
                   selected = "Neighborhood"),
       
       # Select the number of smoothing splines vs. SalePrice
       sliderInput(inputId = "splines",
                   label = "Smoothing splines:", 
                   min = 1, max = 19, step = 2,
                   value = 5),
       
       # Select which neighbourhoods to exclude from the plot, as it can get cluttered quickly
       selectInput(inputId = "neighbourhoods",
                   label = "Exclude neighbourhoods:",
                   choices = all_neighbourhoods,
                   selected = exclude_neighbourhoods,
                   multiple = TRUE,
                   selectize = TRUE),
       
       # Place the boxplot on the lower left side
       plotOutput(outputId = "boxPlot")
    ),
    
    # Show scatter plot and table of selected data on the right side (main area)
    mainPanel(
       plotOutput(outputId = "scatterPlot", brush = "plot_brush"),
       htmlOutput(outputId = "instructions"),
       dataTableOutput(outputId = "houseTable"),
       br()
    )
  )
)

# Define server logic required to draw the plots
function(input, output) {
output$scatterPlot <- renderPlot({
  ggplot(data = house %>% filter(!(Neighborhood %in% input$neighbourhoods)),
          aes_string(x = input$x, y = "SalePrice")) +
    geom_point(aes_string(col = input$z)) +
    geom_smooth(method = "lm", formula = y ~ splines::bs(x, input$splines), se = F) +
    labs(
      y = "Sale Price ($)"
    )
})
  
output$boxPlot <- renderPlot({
  ggplot(data = house %>% filter(!(Neighborhood %in% input$neighbourhoods))) +
    geom_boxplot(aes_string(x = input$z, y = "SalePrice", fill = input$z)) +
    theme(legend.position="none",
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    labs(
      x = ""
    )
})
  
output$instructions <- renderUI({
    HTML("Click and drag selection of observations in scatter plot to inspect closer.<br>")
})
  
  # Create data table
output$houseTable <- DT::renderDataTable({
    brushedPoints(house, brush = input$plot_brush) %>% 
      filter(!(Neighborhood %in% input$neighbourhoods)) %>%
      select(SalePrice, Neighborhood, TotalSF, HouseAge, SaleCondition, GarageCars, GarageAge)
})
}
```

```{r echo=FALSE}
library(shiny)
library(DT)
library(readr)
library(dplyr)
library(ggplot2)

# Load data
house <- read.csv('houseprices.csv')

# Get lists of all neighbourhoods and which to exclude by default, for plotting
house$Neighborhood = as.factor(house$Neighborhood) # for levels to be pulled, it needs to be formatted as factor
all_neighbourhoods <- levels(pull(house, 'Neighborhood'))
exclude_neighbourhoods <- setdiff(levels(house$Neighborhood), c("Blmngtn", "StoneBr", "Mitchel", "OldTown", "NoRidge"))

# Define UI for application
fluidPage(
  # Application title
  titlePanel("Iowa Housing: Sale Price"),
  
  # Sidebar with various Exploratory Data Analysis choices
  sidebarLayout(
    sidebarPanel(
       # Select variable for x-axis
       selectInput(inputId = "x",
                   label = "X-axis:",
                   choices = c("GrLivArea", "TotalSF", "TotalBsmtSF", "HouseAge"),
                   selected = "GrLivArea"),
       
       # Select variable for colour
       selectInput(inputId = "z",
                   label = "Colour:",
                   choices = c("Neighborhood", "CentralAir", "MSSubClass", "SaleType", "SaleCondition", "Foundation"),
                   selected = "Neighborhood"),
       
       # Select the number of smoothing splines vs. SalePrice
       sliderInput(inputId = "splines",
                   label = "Smoothing splines:", 
                   min = 1, max = 19, step = 2,
                   value = 5),
       
       # Select which neighbourhoods to exclude from the plot, as it can get cluttered quickly
       selectInput(inputId = "neighbourhoods",
                   label = "Exclude neighbourhoods:",
                   choices = all_neighbourhoods,
                   selected = exclude_neighbourhoods,
                   multiple = TRUE,
                   selectize = TRUE),
       
       # Place the boxplot on the lower left side
       plotOutput(outputId = "boxPlot")
    ),
    
    # Show scatter plot and table of selected data on the right side (main area)
    mainPanel(
       plotOutput(outputId = "scatterPlot", brush = "plot_brush"),
       htmlOutput(outputId = "instructions"),
       dataTableOutput(outputId = "houseTable"),
       br()
    )
  )
)

# Define server logic required to draw the plots
output$scatterPlot <- renderPlot({
  ggplot(data = house %>% filter(!(Neighborhood %in% input$neighbourhoods)),
          aes_string(x = input$x, y = "SalePrice")) +
    geom_point(aes_string(col = input$z)) +
    geom_smooth(method = "lm", formula = y ~ splines::bs(x, input$splines), se = F) +
    labs(
      y = "Sale Price ($)"
    )
})
  
output$boxPlot <- renderPlot({
  ggplot(data = house %>% filter(!(Neighborhood %in% input$neighbourhoods))) +
    geom_boxplot(aes_string(x = input$z, y = "SalePrice", fill = input$z)) +
    theme(legend.position="none",
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    labs(
      x = ""
    )
})
  
output$instructions <- renderUI({
    HTML("Click and drag selection of observations in scatter plot to inspect closer.<br>")
})
  
  # Create data table
output$houseTable <- DT::renderDataTable({
    brushedPoints(house, brush = input$plot_brush) %>% 
      filter(!(Neighborhood %in% input$neighbourhoods)) %>%
      select(SalePrice, Neighborhood, TotalSF, HouseAge, SaleCondition, GarageCars, GarageAge)
})

```


## 2.2 Target audience for the dashboard

# 3. Threshold analysis in a classification model
Interactive visualizations can be useful when working with machine learning classifications models. To illustrate this we have used the "Bank Marketing" dataset from the UCI Machine Learning Repository. The data is related to a direct marketing campaign for a Portuguese bank. Our goal is to predict the outcome of a sales call when a sales representative is offering a new product.

A confusion matrix is used to evaluate the model. The matrix shows when the model gives true or false positives and true or false negatives. Depending on the business case, different measures can be used to evaluate the model. The total accuracy, the share of correct predictions, might not be the adequate metric for a model. _Precision_ and _recall_ needs to be taken into account as well. Precision measures share of correct predictions among the predicted successes, while recall measures the correct predictions among the successful outcomes. Looking at the graph, we see that there is a trade-off between the two.

When the probability of belonging to a class reaches a set threshold, it will be classified to that class. Consequently, it is essential to adjust the threshold and see how the metrics changes. 

```{r echo=TRUE, results='hide'}
library(shiny)
library(ggplot2)
load(file = "logfit.Rdata")

fluidPage(
  
  # Application title
  titlePanel("Threshold Analysis"),
  
  # Sidebar with a slider input for number of bins 
  sidebarLayout(
    sidebarPanel(
      sliderInput("threshold",
                  "Threshold:",
                  min = 0.00,
                  max = 0.50,
                  value = 0.25),
      selectInput("models",
                  "Model",
                  choices = c("log", "lda", "qda"))
    ),
    # Show a plot of the generated distribution
    mainPanel(
      plotOutput("logfit", width = "100%", height = "500px"),
      plotOutput("threshold", width = "100%", height = "500px"),
      width = 12)
  )
)

output$logfit <- renderPlot({
  plot.list[[input$models]] + geom_vline(xintercept=input$threshold, linetype = "dashed", color = "black")
  })
  
output$threshold <- renderPlot({
  confusion.matrix(table(ifelse(as.numeric(threshold.list[[input$models]] > input$threshold), 1, 0), test.data$y),
                   tit = paste0(input$models, " classification probability threshold ", input$threshold * 100, "%"))
  })
```

```{r echo=FALSE}
load(file = "logfit.Rdata")

fluidPage(
  
  # Application title
  titlePanel("Threshold Analysis"),
  
  # Sidebar with a slider input for number of bins 
  sidebarLayout(
    sidebarPanel(
      sliderInput("threshold",
                  "Threshold:",
                  min = 0.00,
                  max = 0.50,
                  value = 0.25),
      selectInput("models",
                  "Model",
                  choices = c("log", "lda", "qda"))
    ),
    # Show a plot of the generated distribution
    mainPanel(
      plotOutput("logfit", width = "100%", height = "500px"),
      plotOutput("threshold", width = "100%", height = "500px"),
      width = 12)
  )
)

 output$logfit<- renderPlot({
    plot.list[[input$models]] + geom_vline(xintercept=input$threshold, linetype = "dashed", color = "black")
  })
  
  output$threshold <- renderPlot({
    confusion.matrix(table(ifelse(as.numeric(threshold.list[[input$models]] > input$threshold), 1, 0), test.data$y),
                     tit = paste0(input$models, " classification probability threshold ", input$threshold * 100, "%"))
    
  })
```
